Analysis

### 2021-01-28 analysis.Rmd
ordinal analysis
comments on Scott's study

```{r}
# get data into shape
d1 <- df %>%
  dplyr::filter(
    !is.na(wb_psychANXIETY),!is.na(age),!is.na(rel_CHURCHFREQ),!is.na(wb_psychMEANING),!is.na(male),!is.na(rel_EXTENTBELIEVEGOD)
  ) %>%
  dplyr::mutate(
    anx = as.numeric(wb_psychANXIETY),
    anx_s = scale(wb_psychANXIETY),
    #put outcome in ST units (see McElreath 2021)  # not necessary
    age_s = scale(age),
    ch = rel_CHURCHFREQ,
    mn  = wb_psychMEANING,
    bg =  rel_EXTENTBELIEVEGOD,
    ch_s = scale(rel_CHURCHFREQ),
    mn_s = scale(wb_psychMEANING),
    bg_s = scale(rel_EXTENTBELIEVEGOD)
  ) %>%
  dplyr::select(anx, ch, mn, bg, anx_s, age_s, ch_s, mn_s, male2, country, male, bg_s)

@@

@@

```{r}
bayestestR::describe_prior(m1)
```

@@

The model.

```{r}
m1a <- brm(
  anx_s ~ age_s  + male + ch_within +  ch_between + (1 + ch_within|country),  
  data = d1,
  family = "gaussian",
  seed = 1234,
  warmup = 1000,
  iter = 3000,
  chains = 4,
  file = "m1_a.rds")

@@

Which we interpret as follows:

@@

```{r cache = TRUE}
plot(ggeffects::ggpredict(
  m1a,
  terms = c("country[all]", "ch_within[minmax]"),
  type = "re",
  ci.lvl = 0.89))  +
  scale_fill_brewer(palette = "Set2") + 
  labs(title = "Predicted effects of church on anxiety: countries X within") + 
  scale_y_continuous(limits = c(-1,1))
```

@@

```{r cache = TRUE}
plot(ggeffects::ggpredict(
  m1a,
  terms = c("ch_between[minmax]"),
  type = "re",
  ci.lvl = 0.89))  +
  scale_fill_brewer(palette = "Set2") + labs(title = "Predicted effects of church on anxiety: between")  + scale_y_continuous(limits = c(-1, 1))
```

@@

```{r}
tidy_ggdag2 <- dagify(
  wb  ~ meaning + ses + male + country  +U2,
  meaning ~ bg +  male + country, 
  bg ~ age + male + country + U1,
  ses ~ country + age + U2,
  exposure =  "bg",
  outcome =   "wb") %>%
  tidy_dagitty()

@@

@@

p<-brms::conditional_effects(m2, "bg_s", 
                              spaghetti = T,
                              nsamples = 300,
                              prob = .89,
                              method = 'fitted')

plot(p, plot = FALSE)[[1]] + scale_y_continuous(limits = c(-1,1))  + xlab("bg_s") + ylab ("anxiety in sd units") + 
  labs(title = "Predicted effects of bg_s")

@@

p<-brms::conditional_effects(m2, "bg_s", 
                              spaghetti = T,
                              nsamples = 300,
                              prob = .89,
                              method = 'fitted')

plot(p, plot = FALSE)[[1]] + scale_y_continuous(limits = c(-1,1))  + xlab("bg_s") + ylab ("anxiety in sd units") + 
  labs(title = "Predicted effects of bg_s")

@@

```{r cache = TRUE}
plot(ggeffects::ggpredict(m2, terms = c("country[all]", "bs_s[minmax]"),type = "re", ci.lvl = 0.89))  +
   scale_fill_brewer(palette= "Set2") + labs(
    title = "Predicted effects of bs_s on anxiety") + 
  scale_y_continuous(limits = c(-1,1))
```

@@

```{r cache = TRUE}
m2b <- brm(
  anx_s ~ age_s  + male  +  bg_within + bg_between +  (1  +  bg_within|country),
  data = d1,
  family = "gaussian",
  seed = 1234,
  warmup = 1000,
  iter = 300,
  chains = 4,
  file = "m2_aa.rds")
model_parameters(m2b, test = "p_direction")
random_parameters(m2b)
```

@@

p<-brms::conditional_effects(m2b, "bg_within", 
                              spaghetti = T,
                              nsamples = 300,
                              prob = .89,
                              method = 'fitted')

plot(p, plot = FALSE)[[1]] + scale_y_continuous(limits = c(-1,1))  + xlab("bg_within") + ylab ("anxiety in sd units") + 
  labs(title = "Predicted effects of bg_within")

@@

p<-brms::conditional_effects(m2b, "bg_between", 
                              spaghetti = T,
                              nsamples = 300,
                              prob = .89,
                              method = 'fitted')

plot(p, plot = FALSE)[[1]] + scale_y_continuous(limits = c(-1,1))  + xlab("bg_between") + ylab ("anxiety in sd units") + 
  labs(title = "Predicted effects of bg_between")

@@

```{r cache = TRUE}
plot(ggeffects::ggpredict(m2b, terms = c("country[all]", "bg_within[minmax]"),type = "re", ci.lvl = 0.89))  +
   scale_fill_brewer(palette= "Set2") + labs(
    title = "Predicted effects of belief on anxiety")
```

To do:

@@

  2. Use stronger priors and include stronger prior predictive checks

  3. Sample from an ordinal outcome distribution: obtain un-transformed data from Suzanne

@@

>>>>

```{r}
# get data into shape
d1 <- df %>%
  dplyr::filter(
    !is.na(wb_psychANXIETY),!is.na(age),!is.na(rel_CHURCHFREQ),!is.na(wb_psychMEANING),!is.na(male),!is.na(rel_EXTENTBELIEVEGOD)
  ) %>%
  dplyr::mutate(
    anx = as.numeric(wb_psychANXIETY),
    anx_s = scale(wb_psychANXIETY),
    #put outcome in ST units (see McElreath 2021)  # not necessary
    age_s = scale(age),
    ch = rel_CHURCHFREQ,
    mn  = wb_psychMEANING,
    bg =  rel_EXTENTBELIEVEGOD,
    ch_s = scale(rel_CHURCHFREQ),
    mn_s = scale(wb_psychMEANING),
    bg_s = scale(rel_EXTENTBELIEVEGOD)
  ) %>%
  dplyr::select(anx, ch, mn, bg, anx_s, age_s, ch_s, mn_s, male2, country, male, bg_s,wb_psychANXIETY)

@@

```{r eval= FALSE}
prior_summary(m1)
```

@@

```{r eval = FALSE}
bayestestR::describe_prior(m1)
```

@@

The model.
wb_psychANXIETY
```{r}
m1a <- brm(
  anx_s ~ age_s  + male + ch_within +  ch_between + (1 + ch_within|country),  
  data = d1,
  family = "gaussian",
  seed = 1234,
  warmup = 1000,
  iter = 3000,
  chains = 4,
  file = "m1_a.rds")

@@

## TRY ORDINAL MODEL (note church attend should be monotonic -- at next iteration)

```{r}
m1o <- brm(
  wb_psychANXIETY ~ ch_s +
    age_s  +
    male +
    (1 + ch_s | country),
  data = d1,
  family = cumulative(link ="logit"),
  seed = 1234,
  warmup = 1000,
  iter = 3000,
  chains = 4,
  file = "m1o"
)
parameters::model_parameters(m1o, test = "p_direction")
parameters::random_parameters(m1o)
```

This resolves the problem
```{r}
# YES
pp_check(m1o)
```

Trace plots:

```{r, cache = TRUE}
mcmc_plot(m1o,
          type = "trace")
```

Model summary
```{r}
parameters::model_parameters(m1o)
```

Plot model
```{r, cache = TRUE}
library(viridis)
p0b<-brms::conditional_effects(m1o, "ch_s", 
                              spaghetti = T,
                              nsamples = 200,
                              categorical = T,
                              prob = .89,
                              method = 'fitted')

plot(p0b, plot = FALSE)[[1]]   + xlab("ch_s") + ylab ("anxiety 1-5") +
  labs(title = "Predicted effects of ch_s")# + scale_y_continuous(limits= c(1,5))# +  scale_color_viridis(discrete=TRUE)

<!-- Another plot method -->

<!-- ```{r, cache = TRUE} -->
<!-- library(viridis) -->
<!-- p0b1<-brms::conditional_effects(m1o, "ch_s",  -->
<!--                               spaghetti = T, -->
<!--                               nsamples = 200, -->
<!--                              categorical = F, -->
<!--                               prob = .89, -->
<!--                               method = 'fitted') -->

<!-- plot(p0b1, plot = FALSE)[[1]]   + xlab("ch_s") + ylab ("anxiety 1-5") + -->
<!--   labs(title = "Predicted effects of ch_s")+ scale_y_continuous(limits= c(1,5))# +  scale_color_viridis(discrete=TRUE)  -->

<!-- ``` -->

### without demeaning

```{r}
m1_o <- brm(
  wb_psychANXIETY ~ age_s  + male + ch_within +  ch_between + (1 + ch_within|country),  
  data = d1,
  family = cumulative(link ="logit"),
  seed = 1234,
  warmup = 1000,
  iter = 3000,
  chains = 4,
  file = "m1a_o")
```

### Within /between demeaning

```{r}
m1a_o <- brm(
  wb_psychANXIETY ~ age_s  + male + ch_within +  ch_between + (1 + ch_within|country),  
  data = d1,
  family = cumulative(link ="logit"),
  seed = 1234,
  warmup = 1000,
  iter = 3000,
  chains = 4,
  file = "m1a_o")
```

This resolves the problem
```{r}
# YES
pp_check(m1a_o)
```

Trace plots:

```{r, cache = TRUE}
mcmc_plot(m1a_o,
          type = "trace")
```

Model summary
```{r}
parameters::model_parameters(m1a_o)
```

Within: NOTHING
```{r, cache = TRUE}
library(viridis)
pch_w<-brms::conditional_effects(m1a_o, "ch_within", 
                              spaghetti = T,
                              nsamples = 200,
                              categorical = TRUE,
                              prob = .89,
                              method = 'fitted')

ppch_w<-plot(pch_w, plot = FALSE)[[1]]    + ylab ("anxiety in sd units") +
  labs(title = "Predicted effects of Church_within") # +  scale_color_viridis(discrete=TRUE) 
ppch_w

BETWEEN: Yes, and at the highest levels of anxiety.

pch_b<-brms::conditional_effects(m1a_o, "ch_between", 
                              spaghetti = T,
                              nsamples = 200,
                              prob = .89,
                              categorical = TRUE,
                              method = 'fitted')

ppch_b<- plot(p, plot = FALSE)[[1]]   + xlab("ch_between") + ylab ("anxiety: 1-5") + 
  labs(title = "Predicted effects of Church_between")
ppch_b
```

<!-- ### Different plot method -->

<!-- Summing over the categories: -->

<!-- ```{r, cache = TRUE} -->
<!-- library(viridis) -->
<!-- p0a<-brms::conditional_effects(m1a_o, "ch_within",  -->
<!--                               spaghetti = T, -->
<!--                               nsamples = 200, -->
<!--                              # categorical = TRUE, -->
<!--                               prob = .89, -->
<!--                               method = 'fitted') -->

<!-- plot(p0a, plot = FALSE)[[1]]   + xlab("bg_within") + ylab ("anxiety in sd units") + -->
<!--   labs(title = "Predicted effects of bg_within")  + scale_y_continuous(limits = c(1,5))# +  scale_color_viridis(discrete=TRUE)  -->
<!-- ``` -->

<!-- ```{r cache = TRUE} -->

<!-- ps<-brms::conditional_effects(m1a_o, "ch_between",  -->
<!--                               spaghetti = T, -->
<!--                               nsamples = 200, -->
<!--                               prob = .89, -->
<!--                              # categorical = TRUE, -->
<!--                               method = 'fitted') -->

<!-- plot(ps, plot = FALSE)[[1]]   + xlab("bg_between") + ylab ("anxiety: 1-5") +  -->
<!--   labs(title = "Predicted effects of bg_between") + scale_y_continuous(limits = c(1,5)) -->

<!-- ``` -->

## Summary of Church attendence Model
Which we interpret as follows:

@@

```{r cache = TRUE}
ppch_w
```

@@

```{r cache = TRUE}
ppch_b
```

@@

```{r}
tidy_ggdag2 <- dagify(
  wb  ~ meaning + ses + male + country  + U2,
  meaning ~ bg +  male + country,
  bg ~ age + male + country + U1,
  ses ~ country + age + U2,
  exposure =  "bg",
  outcome =   "wb"
) %>%
  tidy_dagitty()

@@

tidy_ggdag2 <- dagify(
  wb  ~ meaning + ses + male + country  + U2,
  meaning ~ bg +  male + country,
  bg ~ age + male + country + U1,
  ses ~ country + age + U2,
  exposure =  "bg",
  outcome =   "wb"
) %>%
  tidy_dagitty()

```{r}

@@

p <- brms::conditional_effects(
  m2,
  "bg_s",
  spaghetti = T,
  nsamples = 300,
  prob = .89,
  method = 'fitted'
)

plot(p, plot = FALSE)[[1]] + scale_y_continuous(limits = c(-1, 1))  + xlab("bg_s") + ylab ("anxiety in sd units") +
  labs(title = "Predicted effects of bg_s")

@@

```{r cache = TRUE}
p <- brms::conditional_effects(
  m2,
  "bg_s",
  spaghetti = T,
  nsamples = 300,
  prob = .89,
  method = 'fitted'
)

plot(p, plot = FALSE)[[1]] + scale_y_continuous(limits = c(-1, 1))  + xlab("bg_s") + ylab ("anxiety in sd units") +
  labs(title = "Predicted effects of bg_s")

@@

```{r cache = TRUE}
plot(ggeffects::ggpredict(
  m2,
  terms = c("country[all]", "bs_s[minmax]"),
  type = "re",
  ci.lvl = 0.89
))  +
  scale_fill_brewer(palette = "Set2") + labs(title = "Predicted effects of bs_s on anxiety") +
  scale_y_continuous(limits = c(-1, 1))
```

@@

```{r cache = TRUE}
m2b <- brm(
  anx_s ~ age_s  + male  +  bg_within + bg_between +  (1  +  bg_within |
                                                         country),
  data = d1,
  family = "gaussian",
  seed = 1234,
  warmup = 1000,
  iter = 300,
  chains = 4,
  file = "m2_aa.rds"
)
model_parameters(m2b, test = "p_direction")
random_parameters(m2b)
```

## Checks

@@

p <- brms::conditional_effects(
  m2b,
  "bg_within",
  spaghetti = T,
  nsamples = 300,
  prob = .89,
  method = 'fitted'
)

plot(p, plot = FALSE)[[1]] + scale_y_continuous(limits = c(-1, 1))  + xlab("bg_within") + ylab ("anxiety in sd units") +
  labs(title = "Predicted effects of bg_within")

@@

```{r cache = TRUE}
p<-brms::conditional_effects(m2b, "bg_between", 
                              spaghetti = T,
                              nsamples = 300,
                              prob = .89,
                              method = 'fitted')

plot(p, plot = FALSE)[[1]] + scale_y_continuous(limits = c(-1, 1))  + xlab("bg_between") + ylab ("anxiety 1-5") +
  labs(title = "Predicted effects of bg_between")

@@

```{r cache = TRUE}
plot(ggeffects::ggpredict(
  m2b,
  terms = c("country[all]", "bg_within[minmax]"),
  type = "re",
  ci.lvl = 0.89
))  +
  scale_fill_brewer(palette = "Set2") + labs(title = "Predicted effects of belief on anxiety")
```

## TRY ORDINAL MODEL

## Without demeaning
```{r}
m2o <- brm(
  wb_psychANXIETY ~ bg_s +
    age_s  +
    male +
    (1 + ch_s | country),
  data = d1,
  family = cumulative(link = "logit"),
  seed = 1234,
  warmup = 1000,
  iter = 3000,
  chains = 4,
  file = "m2o"
)
parameters::model_parameters(m2o, test = "p_direction")
parameters::random_parameters(m2o)
```

This resolves the problem
```{r}
# YES
pp_check(m2o)
```

Trace plots:

```{r, cache = TRUE}
mcmc_plot(m2o,
          type = "trace")
```

Model summary
```{r}
parameters::model_parameters(m2o)
```

Plot model: nothing

```{r, cache = TRUE}
library(viridis)
p0 <- brms::conditional_effects(
  m2o,
  "bg_s",
  spaghetti = T,
  categorical = TRUE,
  nsamples = 200,
  prob = .89,
  method = 'fitted'
)

plot(p0, plot = FALSE)[[1]]   + xlab("bg") + ylab ("anxiety 1-5") +
  labs(title = "Predicted effects of bg") # +  scale_color_viridis(discrete=TRUE) 
```

###  With demeaning

```{r}
m2_o <- brm(
  wb_psychANXIETY ~ age_s  + male + bg_within +  bg_between + (1 + ch_within|country),  
  data = d1,
  family = cumulative(link ="logit"),
  seed = 1234,
  warmup = 1000,
  iter = 3000,
  chains = 4,
  file = "m2_o")
```

This resolves the problem
```{r}
# YES
pp_check(m2_o)
```

Trace plots:

```{r, cache = TRUE}
mcmc_plot(m2_o,
          type = "trace")
```

Model summary
```{r}
parameters::model_parameters(m2_o)
```

Within: NOTHING
```{r, cache = TRUE}
library(viridis)
p0bg <- brms::conditional_effects(
  m2_o,
  "bg_within",
  spaghetti = T,
  nsamples = 200,
  categorical = TRUE,
  prob = .89,
  method = 'fitted'
)

ppbg_w<- plot(p0bg, plot = FALSE)[[1]]   + xlab("bg_within") + ylab ("anxiety 1-5") +
  labs(title = "Predicted effects of BeliefGod_within") # +  scale_color_viridis(discrete=TRUE) 
ppbg_w
```

BETWEEN: Yes, and at the highest levels of anxiety.

```{r cache = TRUE}
pbg <- brms::conditional_effects(
  m2_o,
  "bg_between",
  spaghetti = T,
  nsamples = 200,
  prob = .89,
  categorical = TRUE,
  method = 'fitted'
)

ppbg_b<- plot(pbg, plot = FALSE)[[1]]   + xlab("bg_between") + ylab ("anxiety: 1-5") +
  labs(title = "Predicted effects of BeliefGod_between")
ppbg_b
```

## Summary graph

```{r}
library(patchwork)
ppch_w +  ppbg_w  + ppch_b + ppbg_b + plot_annotation("What is the relationship between belief church and belief on anxiety?", subtitle = "No evidence for within-person differences, evidence for between country differences at the high end of anxiety", tag_levels = "a") +  plot_layout(guides = 'collect')

#bulbulia_summary_graph_many_analysts
```

## To do

@@

  2. Remove clutter.

  3. Use stronger priors and include stronger prior predictive checks

  3. Sample from an ordinal outcome distribution: obtain un-transformed data from Suzanne NOTE: DONE!, but still need to get church frequency as a monotonic predictor!

@@

## APPENDIX
